---
title: "BSHES_740_Assignment_1"
author: "Akshay Deverakonda"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#Installing  and getting set up

library(pacman)
pacman::p_load(rtweet, tm,tidyverse, wordcloud,SnowballC, lsa, stringr, RWeka, wordcloud2, tidytext, quanteda, quanteda.textstats)


setwd("C:/Users/dever/OneDrive/Documents/Emory/Classes/Semester 3 - Fall 2024/BSHES 740")



```

1. What are the date ranges for the two sets? What information is provided in the CSV files?
What are the languages in which tweets have been posted? (2 points)

```{r}

TADA1 <- read.csv("TADA_A1_set1.csv")
head(TADA1)


TADA2 <- read.csv("TADA_A1_set2.csv")
head(TADA2)


#Date ranges for TADA 1
first_1 <- TADA1 %>% 
  arrange(date) %>%
  head(1) %>%
  pull(date)


last_1 <- TADA1 %>% 
  arrange(date) %>%
  tail(1) %>%
  pull(date)

print(first_1)
print(last_1)

first_2 <- TADA2 %>% 
  arrange(date) %>%
  head(1) %>%
  pull(date)
 

last_2 <- TADA2 %>% 
  arrange(date) %>%
  tail(2) %>%
  pull(date)

print(first_2)
print(last_2)

lang1 <- unique(TADA1$lang)

lang2 <- unique(TADA2$lang)

lang_union <- union(lang1,lang2)
lang_union
length(lang_union)

```
For "TADA_A1_set1", the data range from November 1st, 2019 to January 30th, 2020
For "TADA_A1_set2", the data range from February 1st, 2020 to April 29th, 2020

For each tweet, the datasets have a unique ID, a date, the language of the tweet, the tweet itself, and a location for the tweet. There are tweets in 40 languages between the two datasets, and also tweets with an "undetermined" language category.


2. What is the total number of posts in set 1? What is the total in set 2? (1 point)

```{r}
nrow(TADA1)
nrow(TADA2)

```
There are 171012 posts in "TADA_A1_set1", and 226852 posts in "TADA_A1_set2"

Setting up corpuses and document term matrices for the two datasets

```{r}

#Merging documents, then creating one corpus from them

TADA_tweets <- rbind (TADA1, TADA2)
class(TADA_tweets)

nrow(TADA_tweets)

#Setting up corpus for TADA combined dataset

TADA_texts <- TADA_tweets$text
TADA_texts_corpus <- Corpus(VectorSource(TADA_texts))

#meta(TADA_texts_corpus, "id") <- TADA_tweets$id
#meta(TADA_texts_corpus, "date") <- TADA_tweets$date



#Pre-processing TADA combined corpus
TADA_texts_corpus <- tm_map(TADA_texts_corpus, content_transformer(tolower))
TADA_texts_corpus <- tm_map(TADA_texts_corpus, removePunctuation)
TADA_texts_corpus <- tm_map(TADA_texts_corpus, removeNumbers)
TADA_texts_corpus  <- tm_map(TADA_texts_corpus , function(x) removeWords(x, stopwords()))
TADA_texts_corpus <- tm_map(TADA_texts_corpus, stemDocument)

#Making document term matrix

tada_dtm <- DocumentTermMatrix(TADA_texts_corpus)
tada_dtm_r <- removeSparseTerms(tada_dtm, sparse = .99)

```
3. How many tweets are there for methadone, suboxone, and fentanyl in total? Tip: sometimes
alternative expressions are used for substances (eg., fent for fentanyl). (2 points)
```{r}

#Looking at combined TADA text file
#length(grep("methadone",TADA_texts))

substance <- "suboxon|Suboxon|fent|Fent|methadon|Methadon"
count <- length(grep(substance,TADA_texts))
print(count)


```
There are 376,266 tweets in total that mention one or more of the three substances.

4. Are there fentanyl analogs that are also being discussed (eg., carfentanil)?

```{r}
analog_c <- "carfentanil|Carfentanil|carfent|Carfent"
analog_f <- "furanylfentanyl|Furanylfentanyl|furanylfent|Furanylfent"
analog_a <- "acetylfentanyl|Acetylfentanyl|acetylfent|Acetylfent"
analog_m <- "methylfentanyl|Methylfentanyl|methylfent|Methylfent"

count_analog_c <- length(grep(analog_c,TADA_texts))
count_analog_f <- length(grep(analog_f,TADA_texts))
count_analog_a <- length(grep(analog_a,TADA_texts))
count_analog_m <- length(grep(analog_m,TADA_texts))


print(count_analog_c)
print(count_analog_f)
print(count_analog_a)
print(count_analog_m)





```
Yes, carfentanil, furanylfentanyl, and acetylfentanyl are all mentioned in the tweets

5. What are some of the topics that are most closely associated with each of the three
substances? The top 5-10 topics (if relevant) are acceptable. (2 points)

```{r}

#Three substances of concern are methadone, suboxone, and fentanyl

#Fentanyl

fentanyl <- as.data.frame(findAssocs(tada_dtm_r, "fentanyl",0.05))

fentanyl_fil <- top_n(fentanyl,10)
fentanyl_fil


fentanyl_fil %>%
  rownames_to_column() %>%
  ggplot(aes(x=reorder(rowname,fentanyl),y=fentanyl)) + geom_point(size=4) + 
  coord_flip() + ylab('Correlation') + xlab('Term') + 
  ggtitle('Terms correlated with Fentanyl')



#Methadone

methadone <- as.data.frame(findAssocs(tada_dtm_r, "methadon",0.05))

methadone_fil <- top_n(methadone,10)
methadone_fil


methadone_fil %>%
  rownames_to_column() %>%
  ggplot(aes(x=reorder(rowname,methadon),y=methadon)) + geom_point(size=4) + 
  coord_flip() + ylab('Correlation') + xlab('Term') + 
  ggtitle('Terms correlated with Methadone')

#Suboxone

suboxone <- as.data.frame(findAssocs(tada_dtm_r, "suboxon",0.05))


suboxone_fil <- top_n(suboxone,10)
suboxone_fil

suboxone_fil %>%
  rownames_to_column() %>%
  ggplot(aes(x=reorder(rowname,suboxon),y=suboxon)) + geom_point(size=4) + 
  coord_flip() + ylab('Correlation') + xlab('Term') + 
  ggtitle('Terms correlated with Suboxone')

```





6 - generating word cloud

```{r}

#Fentanyl wordcloud
wordcloud(row.names(fentanyl_fil),fentanyl_fil$fentanyl, min.freq=10, max.words=30, scale=c(4,.3), random.order=FALSE, colors=brewer.pal(12, "Set3"))

#Methadone wordcloud
wordcloud(row.names(methadone_fil),methadone_fil$methadon, min.freq=0.09, max.words=30, scale=c(4,.3), random.order=FALSE, colors=brewer.pal(12, "Set3"))


#Suboxone wordcloud
wordcloud(row.names(suboxone_fil),suboxone_fil$suboxon, min.freq=0.08, max.words=30, scale=c(4,.3), random.order=FALSE, colors=brewer.pal(12, "Set3"))


```



7. Generate appropriate time-series figures to compare how the frequencies of mentions of
these substances differ. (2 points)

```{r}

TADA_df <- data.frame(text = TADA_tweets$text,
                      date = TADA_tweets$date)


freq <- TADA_df %>%
  corpus(text_field="text") %>%
  tokens(remove_numbers = TRUE, remove_punct = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_wordstem(language = quanteda_options("language_stemmer")) %>%
  dfm() %>%
  textstat_frequency(groups = date) 
 

#Fentanyl
freq_f <- freq %>%
   filter(feature == "fentanyl") %>%
  select(frequency, group)

freq_f$group <- as.Date(freq_f$group)

ggplot(data=freq_f, aes(x=group, y=frequency)) +
  geom_line() + ggtitle("Frequency (counts) of mentions of fentanyl over time") + 
  xlab("Months (November 2019 - April 2020)") + ylab("Number of mentions") +
geom_vline(xintercept = as.numeric(as.Date("2020-03-13")), 
             linetype = "dashed", color = "black") +
  annotate("text", x = as.Date("2020-03-13"), 
           y = max(freq_f$frequency, na.rm = TRUE), 
           label = "March 13, 2020", 
           vjust = -0.5, hjust = 1.1)

#Methadone

freq_m <- freq %>%
   filter(feature == "methadon") %>%
  select(frequency, group)

freq_m$group <- as.Date(freq_m$group)

ggplot(data=freq_m, aes(x=group, y=frequency)) +
  geom_line() + ggtitle("Frequency (counts) of mentions of methadone over time")  + 
  xlab("Months (November 2019 - April 2020)") + ylab("Number of mentions") +
geom_vline(xintercept = as.numeric(as.Date("2020-03-13")), 
             linetype = "dashed", color = "black") +
  annotate("text", x = as.Date("2020-03-13"), 
           y = max(freq_m$frequency, na.rm = TRUE), 
           label = "March 13, 2020", 
           vjust = -0.5, hjust = 1.1)

#Suboxone

freq_s <- freq %>%
   filter(feature == "suboxon") %>%
  select(frequency, group)

freq_s$group <- as.Date(freq_s$group)


ggplot(data=freq_s, aes(x=group, y=frequency)) +
  geom_line() + ggtitle("Frequency (counts) of mentions of suboxone over time")  + 
  xlab("Months (November 2019 - April 2020)") + ylab("Number of mentions") +
geom_vline(xintercept = as.numeric(as.Date("2020-03-13")), 
             linetype = "dashed", color = "black") +
  annotate("text", x = as.Date("2020-03-13"), 
           y = max(freq_s$frequency, na.rm = TRUE), 
           label = "March 13, 2020", 
           vjust = -0.5, hjust = 1.1)


```


8. Find the top 10 most frequent bigrams in each of the three sets. Plot a bar chart for these. (2
points)’

```{r}

#Tokenizing
TADA_text <- tokens(TADA_tweets$text,remove_numbers = TRUE, remove_punct = TRUE)


#Lowercasing
TADA_text <- tokens_tolower(TADA_text)

#Removing stopwords
TADA_text <- tokens_remove(TADA_text, stopwords("english"))

#Stemming
TADA_text <- tokens_wordstem(TADA_text,language = quanteda_options("language_stemmer") )

#Creating bigrams
TADA_bigrams <- tokens_ngrams(TADA_text, n = 2:2)
print(TADA_bigrams)

TADA_dfm <- dfm(TADA_bigrams)
head(TADA_dfm)
#TADA_dfm <- dfm_trim(TADA_dfm, min_docfreq = 0.01)



#Filtering for fentanyl-related bigrams
TADA_dfm_f <- dfm_select(TADA_dfm,pattern="fentanyl",selection="keep",valuetype="regex",case_insensitive="TRUE")
print(TADA_dfm_f)

test_f <- as.data.frame(topfeatures(TADA_dfm_f,10))
colnames(test_f) <- c("Count")
test_f

ggplot(data=test_f, aes(x = reorder(row.names(test_f), -Count), y=Count)) +
  geom_bar(stat="identity", width = 0.7) + 
    theme(axis.text.x = element_text(angle = 90)) +
  xlab("Bigram term") + ylab("Frequency") +
  ggtitle("Top 10 bigrams of fentanyl-related terms across both datasets")

#Filtering for methadone-related bigrams
TADA_dfm_m <- dfm_select(TADA_dfm,pattern="methadon",selection="keep",valuetype="regex",case_insensitive="TRUE")

test_m <- as.data.frame(topfeatures(TADA_dfm_m,10))
colnames(test_m) <- c("Count")


ggplot(data=test_m, aes(x = reorder(row.names(test_m), -Count), y=Count)) +
  geom_bar(stat="identity", width = 0.7) + 
    theme(axis.text.x = element_text(angle = 90)) +
  xlab("Bigram term") + ylab("Frequency") +
  ggtitle("Top 10 bigrams of methadone-related terms across both datasets")
  

#Filtering for suboxone related bigrams

TADA_dfm_s <- dfm_select(TADA_dfm,pattern="suboxon",selection="keep",valuetype="regex",case_insensitive="TRUE")

test_s <- as.data.frame(topfeatures(TADA_dfm_s,10))
colnames(test_s) <- c("Count")


ggplot(data=test_s, aes(x = reorder(row.names(test_s), -Count), y=Count)) +
  geom_bar(stat="identity", width = 0.7) + 
    theme(axis.text.x = element_text(angle = 90)) +
  xlab("Bigram term") + ylab("Frequency") +
  ggtitle("Top 10 bigrams of suboxone-related terms across both datasets")
  


```

